BRiM: A Modular Bicycle-Rider Modeling Framework

### 2023-12-23 paper.tex

In the trajectory tracking optimization problems, the task is to follow a target path representing a lane change.
The path is defined by the equation
\begin{equation}
    \bar{q}_2\left(q_1\right) = \begin{cases}
        0 \, , & \hphantom{X_1 - X_S \leq } q_1 < X_s \\ 
        \frac{1}{2} X_2\left(1 - \cos{\left(\frac{q_1 - X_s}{X_1 - 2X_s}\pi\right)}\right) \, , & \hphantom{X_1 - {}} X_s \leq q_1 \leq X_1 - X_s \\
        X_2 \, , & X_1 - X_s < q_1 \\
    \end{cases} \, ,
\end{equation}
where $q_1$ and $q_2$ are the states describing the position of the bicycle in the $x$ (longitudinal) and $y$ (lateral) directions of the Newtonian frame respectively, $\bar{x}_2$ is the target lateral position as a function of $x_1$, $X_1$ and $X_2$ are the longitudinal length and lateral displacement of the target path respectively, and $X_s$ is the length of straight at the beginning and end of the path.

@@

We add constraints to define the initial and final state of the system's motion.
In optimizations \#\numrange[range-phrase=--]{1}{3} we enforce
\begin{equation*}
    \begin{aligned}
        q_1\left(t_0\right) &= 0.0 \\
        q_2\left(t_0\right) &= 0.0 \\
        q_3\left(t_0\right) &= 0.0 \\
        q_4\left(t_0\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_1\left(t_F\right) - X_1 &= 0.0 \\
        q_2\left(t_F\right) - X_2 &= 0.0 \\
        q_3\left(t_F\right) &= 0.0 \\
        q_4\left(t_F\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_6\left(t_0\right) &= 0.0 \\
        q_7\left(t_0\right) &= 0.0 \\
        q_8\left(t_0\right) &= 0.0 \\
        \\
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \\
        q_7\left(t_F\right) &= 0.0 \\
        \\
        \\
    \end{aligned}
\end{equation*}
For the remaining generalized coordinate ($q_4$) and generalized speeds ($u_i, i \in \mathbb{N} \cap \left[1,8\right]$) we enforce periodicity constraints between the values at $t_0$ and $t_F$ if the variable acts along the longitudinal axis.
Otherwise, we enforce that the value at $t_0$ is the negative of the value at $t_F$ as this results in the solution being symmetric about the longitudinal axis.
Additionally, in optimization \#4 the generalized coordinate and generalized speed for the suspension are constrained at $t_0$ such that the suspension compression and weight through the handlebars are in equilibrium.

@@

We solved each of the six trajectory tracking optimization problems using \opty{}~\citep{moore_2018_opty}, a Python package for solving optimization problems involving dynamic systems using direct collocation~\citep{betts_2010_practical}. \opty{} is designed with \sympy{} in mind and thus interfaces seamlessly with \brim{}. \opty{} transcribes the optimization problem into a nonlinear programming problem, which is then solved numerically using the interior-point solver \ipopt{}~\citep{biegler_2009_large}. Each optimization problem was transcribed with 100 nodes and the dynamics were collocated using the backward Euler method. \ipopt{}'s default settings were used with the exception of enabling gradient-based scaling.

@@

\begin{figure}
    \centering
	\def\svgwidth{\textwidth}
    \input{figures/torques_all.pdf_tex}
    \caption{Comparison of the optimal inputs for optimizations \#\numrange[range-phrase=--]{1}{6}.}
    \label{fig:torques_all}
\end{figure}

>>>>

In the trajectory tracking optimization problems, the task is to follow a target path representing a lane change.
The path is defined by the equation
\begin{equation}
    \bar{q}_2\left(q_1\right) = \begin{cases}
        0 \, , & \hphantom{X_1 - X_S \leq } q_1 < X_s \\ 
        \frac{1}{2} X_2\left(1 - \cos{\left(\frac{q_1 - X_s}{X_1 - 2X_s}\pi\right)}\right) \, , & \hphantom{X_1 - {}} X_s \leq q_1 \leq X_1 - X_s \\
        X_2 \, , & X_1 - X_s < q_1 \\
    \end{cases} \, ,
\end{equation}
where $q_1$ and $q_2$ are the states describing the position of the bicycle in the $x$ (longitudinal) and $y$ (lateral) directions of the Newtonian frame respectively, $\bar{q}_2$ is the target lateral position as a function of $q_1$, $X_1$ and $X_2$ are the longitudinal length and lateral displacement of the target path respectively, and $X_s$ is the length of straight at the beginning and end of the path.

@@

We add constraints defining the initial and final state of the system's motion to be in the nominal configuration.
In optimizations \#\numrange[range-phrase=--]{1}{6} we enforce
\begin{equation*}
    \begin{aligned}
        q_1\left(t_0\right) &= 0.0 \\
        q_2\left(t_0\right) &= 0.0 \\
        q_3\left(t_0\right) &= 0.0 \\
        q_4\left(t_0\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_1\left(t_F\right) - X_1 &= 0.0 \\
        q_2\left(t_F\right) - X_2 &= 0.0 \\
        q_3\left(t_F\right) &= 0.0 \\
        q_4\left(t_F\right) &= 0.0
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        q_6\left(t_0\right) &= 0.0 \\
        q_7\left(t_0\right) &= 0.0 \\
        q_8\left(t_0\right) &= 0.0 \\
        \\
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \\
        q_7\left(t_F\right) &= 0.0 \\
        \\
        \\
    \end{aligned}
\end{equation*}

The velocity at the end of the path is enforced to be the same as at the start using periodic constraints on the angular velocity of both wheels, $u_6$ and $u_8$.
Additionally, in optimization \#4 the generalized coordinate and generalized speed for the suspension are constrained at $t_0$ such that the suspension compression and weight through the handlebars are in equilibrium.

@@

We solved each of the six trajectory tracking optimization problems using \opty{}~\citep{moore_2018_opty}, a Python package for solving optimization problems involving dynamic systems using direct collocation~\citep{betts_2010_practical}. \opty{} is designed with \sympy{} in mind and thus interfaces seamlessly with \brim{}. \opty{} transcribes the optimization problem into a nonlinear programming problem, which is then solved numerically using the interior-point solver \ipopt{}~\citep{biegler_2009_large}. Each optimization problem was transcribed with 100 nodes, resulting in a sample rate of \SI{50}{\hertz}, and the dynamics were collocated using the backward Euler method. \ipopt{}'s default settings were used with the exception of enabling gradient-based scaling.

@@

\begin{figure}
    \centering
	\def\svgwidth{\textwidth}
    \input{figures/torques_all.pdf_tex}
    \caption{Comparison of the optimal inputs for optimizations \#\numrange[range-phrase=--]{1}{6}. Note that \#6 is missing in the steering torque graph because elbow torques are used.}
    \label{fig:torques_all}
\end{figure}
